Local to global N input modelling

### 2025-06-13 manuscript.qmd
Bit of cleanup

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")
```

@@

source(here("00_model_functions.R"))

# Input paths
input_farm_coords_path <- here() %>% file.path("data", "_general_data", "farm_locations")
input_farm_sst_path <- here() %>% file.path("data", "_general_data", "SST")
input_species_param_path <- here() %>% file.path("data", "atlantic_salmon", "params")
input_feed_profile_path <- here() %>% file.path("data", "_general_data", "diets")

# Output paths
output_farm_data_path <- here() %>% file.path("outputs", "farm_data")
output_species_data_path <- here() %>% file.path("outputs", "species_data")
output_sens_data_path <- here() %>% file.path("outputs", "sensitivity_data")
output_farm_growth_data_path <- here() %>% file.path("outputs", "farm_growth_data")
output_cohort_growth_data_path <- here() %>% file.path("outputs", "cohort_growth_data")
data_analysis_path <- here() %>% file.path("outputs", "data_analysis")

# Filenames
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
sens_results_files <- file.path(output_sens_data_path) %>% str_subset("results")

@@

sorted_countries <- farm_coords %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  group_by(country) %>% 
  reframe(av_lat = mean(lat)) %>% 
  arrange(-av_lat) %>% 
  pull(country) %>% 
  unique()
```

@@

p1 <- ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()

@@

(p2 + p3) / p1 + plot_layout(heights = c(1, 1.5))
```

@@

@tbl-params shows the parameter values used to model the growth of farmed Atlantic salmon.

| Parameter | Value | Units | Description | Reference | 
|:-------:|:----:|:----:|:----------------|:-------------------------|
| $\alpha$             | `r unname(params['alpha'])`      | -                        | Feeding catabolism coefficient | *Salvelinus alpinus*, *Oncorhynchus mykiss* (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | J g$^{-1}$               | Energy content of protein      | @baldan_r_2018 | 
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | J g$^{-1}$               | Energy content of lipid        | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | J g$^{-1}$               | Energy content of carbohydrate | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | J g O$_2^{-1}$           | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 |
| pk                   | `r unname(params['pk'])`         | d$^{-1}$                 | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | -                        | Fasting catabolism at 0 $^{^\circ}$C | @macnaughton_using_2019 |
| m                    | `r unname(params['m'])`          | -                        | Weight exponent for anabolism | Experimentally-derived alue over a range of *Oncorhynchus*, *Perca* and *Salmo* species (temps: 8–13$^{\circ}$C) [@broekhuizen_modelling_1994] | 
| n                    | `r unname(params['n'])`          | -                        | Weight exponent for catabolism | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | -                        | Shape coefficient for the feeding function | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$\^{\circ}$C. Coefficient for mean daily energy intake of the moderately fast growing group, 0.4398 (mean fast group), 0.2214 (mean slow group) [@jonsson_thermal_2001] | 
| $T_{max}$            | `r unname(params['Tma'])`        | $^{\circ}$C              | Maximum lethal temperature | Juvenile *Oncorhynchus tshawytscha* (19.8 $\pm$ 0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) [@poletto_unusual_2017] |
| $T_{opt}$            | `r unname(params['Toa'])`        | $^{\circ}$C              | Optimal temperature | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{^\circ}$C [@jonsson_thermal_2001]. Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0 $\pm$ 14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$            | `r unname(params['Taa'])`        | $^{\circ}$C              | Lowest feeding temperature | Fish essentially stop feeding when below 2$^{\circ}$C [@vadboncoeur_lowering_2023] | 
| $\omega$             | `r unname(params['omega'])`      | g O$_2$ g$^{-1}$         | Oxygen consumption:weight loss ratio | Calibrated value from @baldan_r_2018 | 
| a                    | `r unname(params['a'])`          | J g tissue$^{-1}$        | Energy content of fish tissue | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| k                    | `r unname(params['k'])`          | -                        | Weight exponent for energy content | Alt value: 1.65175 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |
| eff                  | `r unname(params['eff'])`        | -                        | Food ingestion efficiency | Range: 95-97% ingestion efficiency [@uglem_does_2020] | 
| $\overline{W}$       | `r unname(params['meanW'])`      | g                        | Dry weight mean | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | g                        | Dry weight standard deviation | Range of 100-150 given so used a mean of 125 and sd of 10 which over 1000 draws from a normal distribution delivers a minimum and maximum close to these numbers [@dempsey_estimating_2023] | 
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | g g$^{-1}$ fish d$^{-1}$ | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\overline{F}$       | `r unname(params['overFmean'])`  | g g$^{-1}$ fish          | Overfeeding rate mean |    |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | g g$^{-1}$ fish          | Overfeeding rate standard deviation |    |
| mort                 | `r unname(params['mortmyt'])`    | d$^{-1}$                 | Natural mortality rate | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon {#tbl-params}

@@

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `terra` [@terra], `targets` [@targets], `future` [@future], `furrr` [@furrr], `renv` [@renv], and the `tidyverse` [@tidyverse].
For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].

>>>>

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(cowplot)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(dtplyr)
library(furrr)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
library(RColorBrewer)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()

remove_unit("g_fish")
remove_unit("kg_fish")
remove_unit("t_fish")
install_unit("g_fish")
install_unit("kg_fish", "1000 g_fish")
install_unit("t_fish", "1000 kg_fish")

packages <- c("magrittr", "dplyr", "tidyr", "stringr", "ggplot2", "arrow", "targets", "sf", "terra", "rnaturalearth", "furrr", "future", "units", "lubridate")

here("00_model_functions.R") %>% source()

remove_unit("gN")
install_unit("gN", "6.25 g") # 1/6.25
```

@@

source(here("00_model_functions.R"))
source(here("00_dirs.R"))

# Filenames
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
sens_results_files <- file.path(output_sens_data_path) %>% list.files(full.names = T) %>% str_subset("results")

@@

sorted_countries <- farm_coords %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  group_by(country) %>% 
  reframe(av_lat = mean(lat)) %>% 
  arrange(-av_lat) %>% 
  pull(country) %>% 
  unique()

farms_geometry <- farms_geometry %>% 
  mutate(country = factor(country, levels = sorted_countries))
```

@@

p_big_map <- ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()

@@

(p2 + p3) / p_big_map + plot_layout(heights = c(1, 1.5))
```

@@

| Parameter | Value | Reference |
|:------:|:------:|:-----------:|
| $\alpha$             | `r unname(params['alpha'])`      | @broekhuizen_modelling_1994 |
| $\epsilon_{P}$       | `r unname(params['epsprot'])`    | @baldan_r_2018 |
| $\epsilon_{L}$       | `r unname(params['epslip'])`     | @baldan_r_2018 |
| $\epsilon_{C}$       | `r unname(params['epscarb'])`    | @baldan_r_2018 |
| $\epsilon_{O_2}$     | `r unname(params['epsO2'])`      | @elliott_energy_1975 |
| $p_k$                | `r unname(params['pk'])`         | @macnaughton_using_2019 |
| $k_0$                | `r unname(params['k0'])`         | @macnaughton_using_2019 |
| $m$                  | `r unname(params['m'])`          | @broekhuizen_modelling_1994 | 
| $n$                  | `r unname(params['n'])`          | @baldan_r_2018 |
| $\beta C$            | `r unname(params['betac'])`      | @jonsson_thermal_2001 |
| $T_{max}$            | `r unname(params['Tma'])`        | @poletto_unusual_2017 | 
| $T_{opt}$            | `r unname(params['Toa'])`        | @jonsson_thermal_2001 | 
| $T_{min}$            | `r unname(params['Taa'])`        | @vadboncoeur_lowering_2023 | 
| $\omega$             | `r unname(params['omega'])`      | @baldan_r_2018 |
| $a$                  | `r unname(params['a'])`          | @van_tien_development_2016 |
| $k$                  | `r unname(params['k'])`          | @van_tien_development_2016 | 
| $eff$                | `r unname(params['eff'])`        | @uglem_does_2020 | 
| $\overline{W}$       | `r unname(params['meanW'])`      | @dempsey_estimating_2023 | 
| $\Delta{W}$          | `r unname(params['deltaW'])`     | @dempsey_estimating_2023 | |
| $\overline{I_{max}}$ | `r unname(params['meanImax'])`   | @lall_nutrition_2009 | 
| $\Delta{I_{max}}$    | `r unname(params['deltaImax'])`  | @lall_nutrition_2009 | 
| $\overline{F}$       | `r unname(params['overFmean'])`  | |
| $\Delta{F}$          | `r unname(params['overFdelta'])` | |
| $mort$               | `r unname(params['mortmyt'])`    | @tvete_towards_2023 |
: Model growth parameters for Atlantic salmon (for double-checking numbers and reference list only) {#tbl-params}

@@

All analysis was conducted in R version `r R.version[['major']]`.`r R.version[['minor']]` "`r R.version[['nickname']]`" [@R_base]. 
Major packages used include `r cite_packages(packages)`.
For a full list of R packages used see the lockfile on [Github](https://www.github.com).
This manuscript was written in Quarto [@quarto] using TinyTex [@tinytex] and the acronyms extension [@acronyms].
